---
title: "p8105_hw1_sjz2120"
date: "2022-09-19"
output: github_document
---

# Prep work
```{r load_packages}
library(tidyverse)
```

# Problem 1
```{r download_penguins_package, eval=FALSE}
install.packages("palmerpenguins")
```

```{r penguins_dataset}
data("penguins", package = "palmerpenguins")

penguins
```

## **Describing the penguins dataset**

* There are 8 variables in this dataset, including: Species, Island, Bill Length (in mm), Bill Depth (in mm), Flipper Length (in mm), Body Mass (in g), Sex, and Year.

* The dataset contains `r nrow(penguins)` rows (observations)
and `r ncol(penguins)` columns (variables).

* The mean flipper length among penguins in this dataset is
`r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.


```{r penguins_flipper_bill_scatterplot}
##Create scatterplot of flipper length (y) vs bill length (x)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper Length (mm)") + ggtitle("Plot of Penguin Flipper vs Bill Length (by species)")

##Save scatterplot to project directory
ggsave("flipper_vs_bill_scatterplot.pdf", height = 4, width = 6)
```


# Problem 2
```{r create_df}
problem2_df = tibble(
  rand_samp_10 = rnorm(10),
  greater_zero = c(rand_samp_10 > 0),
  char_vec_10 = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor_10 = factor(c("high", "medium", "low", "medium","high", "high", "low", "low", "medium", "high"))
)
```


```{r try_to_take_means}
mean(pull(problem2_df, rand_samp_10))
mean(pull(problem2_df, greater_zero))
mean(pull(problem2_df, char_vec_10))
mean(pull(problem2_df, vec_factor_10))
```
We're able to take the mean of variables **rand_samp_10** and **greater_zero** because they are **numeric and logical variable types** (respectively), meaning that we can take the mean of them.

However, we're unable to take the mean of **variables char_vec_10** and **vec_factor_10** because they are character and factor variable types (respectively). Since neither of them are numeric or logical variable types, we can't take the mean of these variables.

```{r apply_as_numeric, results='hide'}
as.numeric(pull(problem2_df, greater_zero))
as.numeric(pull(problem2_df, char_vec_10))
as.numeric(pull(problem2_df, vec_factor_10))
```
For the logical variable **greater_zero,** the FALSE values are recoded as 0 while the TRUE values are recoded as 1 (thus the variable becomes numeric).

For the character variable **char_vec_10,** all of the values become **NA** because there is no way to recode the character values numerically.

For the factor variable **vec_factor_10,** the three different levels are recoded as numbers. In this case, "high" was recoded as 1, "medium" was recoded as 3, and "low" was recoded as 2.

This helps explain what happens when we try to take the mean of each of these non-numeric variables. For the logical variable **greater_zero,** we're able to take the mean because FALSE=0 and TRUE=1 (they can already be considered numbers).

However, for **char_vec_10** and **vec_factor_10** we are  unable to take their means because their values are not numeric; and only **vec_factor_10** can be converted to a numeric variable after using as.numeric.
