---
title: "p8105_hw1_sjz2120"
author: "Sophie Zhang"
date: "2022-09-19"
output: github_document
---

# Prep Work: Loading packages
```{r load_packages}
library(tidyverse)
```

# Problem 1
## Loading the `penguins` dataset
This problem focuses the use of inline R code, plotting, and the behavior of ggplot for variables of different types.

Use the code below to download the a package containing the penguins dataset, but only run it in the console (**code shown below for reference, but using eval=FALSE in the code chunk so that the code will only be displayed, NOT executed**)
```{r download_penguins_package, eval=FALSE}
install.packages("palmerpenguins")
```

Next, use the following code to load the `penguins` dataset:
```{r penguins_dataset}
data("penguins", package = "palmerpenguins")
```

## **Describing the penguins dataset**
```{r}
str(penguins)
penguins
```
Write a short description of the `penguins` dataset (not the `penguins_raw` dataset) using inline R code.

* There are a total of **344 rows/entries** in the `penguins` dataset and a **total of 8 variables**:
    * **3 factor variables:** `Species`, `Island`, `Sex`
    * **5 numeric variables:** `Bill Length (in mm)`, `Bill Depth (in mm)`, `Flipper Length (in mm)`, `Body Mass (in g)`, and `Year`.

* The dataset contains **`r nrow(penguins)` rows (observations)**
and **`r ncol(penguins)` columns (variables)**.

* The mean flipper length among penguins in this dataset is
**`r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm**.

## Creating a scatterplot for `penguins' dataset
Make a scatterplot of flipper_length_mm (y) vs bill_length_mm (x), using color points for the species variable:
```{r penguins_flipper_bill_scatterplot}
##Create scatterplot of flipper length (y) vs bill length (x)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper Length (mm)") + ggtitle("Plot of Penguin Flipper vs Bill Length (by species)")
```

## Exporting the `penguins` flipper vs bill length scatterplot
Export the scatterplot to the project directory using ggsave:
```{r}
##Save scatterplot to project directory
ggsave("flipper_vs_bill_scatterplot.pdf", height = 4, width = 6)
```


# Problem 2
## Creating a dataframe
Create a data frame comprised of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor “levels”

```{r create_df}
problem2_df = tibble(
  var_snorm = rnorm(10, sd = 1),
  var_logic = c(var_snorm > 0),
  var_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  var_factor = factor(c("high", "medium", "low", "medium","high", "high", "low", "low", "medium", "high"))
)
```

## Taking the means
Try to take the mean of each variable in your dataframe. What works and what doesn’t?
```{r try_to_take_means}
mean(pull(problem2_df, var_snorm))
mean(pull(problem2_df, var_logic))
mean(pull(problem2_df, var_char))
mean(pull(problem2_df, var_factor))
```
We're able to take the mean of variables `var_snorm` and `var_logic` because they are **numeric and logical variable types** (respectively), meaning that their values can be represented numerically. For instance, when taking the mean for logical variable types they are treated as TRUE=1 and FALSE=0, and we can take the mean of these 1s and 0s.

However, we're unable to take the mean of variables `var_char` and `var_factor` because they are character and factor variable types (respectively). Since neither of them in their current forms are able to be represented numerically, we can't take the mean of these variables.

## Applying `as.numeric`
```{r apply_as_numeric, results='hide'}
as.numeric(pull(problem2_df, var_logic))
as.numeric(pull(problem2_df, var_char))
as.numeric(pull(problem2_df, var_factor))
```
For the logical variable `var_logic`, the FALSE values are recoded as 0 while the TRUE values are recoded as 1 (thus the variable becomes numeric).

For the character variable `var_char`, all of the values become **NA** because there is no way to recode the character values numerically.

For the factor variable `var_factor`, the three different levels are recoded as numbers. In this case, "high" was recoded as 1, "medium" was recoded as 3, and "low" was recoded as 2.


**Does `as.numeric` help explain what happens when you try to take the mean?**

This helps explain what happens when we try to take the mean of each of these non-numeric variables. For the logical variable `var_logic`, we're able to take the mean because R considers FALSE=0 and TRUE=1. However, for `var_char` and `var_factor` we are  unable to take their means because their values are not numeric; and only `var_factor` can be converted to a numeric variable after using as.numeric.
