---
title: "p8105_hw1_sjz2120"
date: "2022-09-19"
output: github_document
---

# Prep Work: Loading packages
```{r load_packages}
library(tidyverse)
```

# Problem 1
```{r download_penguins_package, eval=FALSE}
install.packages("palmerpenguins")
```

```{r penguins_dataset}
data("penguins", package = "palmerpenguins")

penguins
```

## **Describing the penguins dataset**

* There are 8 variables in this dataset, including: Species, Island, Bill Length (in mm), Bill Depth (in mm), Flipper Length (in mm), Body Mass (in g), Sex, and Year.

* The dataset contains `r nrow(penguins)` rows (observations)
and `r ncol(penguins)` columns (variables).

* The mean flipper length among penguins in this dataset is
`r mean(pull(penguins, flipper_length_mm), na.rm = TRUE)` mm.


```{r penguins_flipper_bill_scatterplot}
##Create scatterplot of flipper length (y) vs bill length (x)
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + geom_point() + xlab("Bill Length (mm)") + ylab("Flipper Length (mm)") + ggtitle("Plot of Penguin Flipper vs Bill Length (by species)")

##Save scatterplot to project directory
ggsave("flipper_vs_bill_scatterplot.pdf", height = 4, width = 6)
```

